Collaborative Visual Analysis on the Web with RCloud

### 2015-03-31 discussion.tex
move hard parts to end of discussion
further tweaking and pruning

{\bf Exploration is hard.}
R is a popular language for exploratory data analysis, and we built RCloud
as a way for data scientists to seamlessly transition from EDA to sharing
and presentation. But the interface we implemented for combining sharing and presentation
into the R command line interferes with the exploration aspect.

In existing R tools, the command prompt is used to try things out. It is very
fast and there is no commitment: nothing gets saved and there is nothing to
later clean up. In contrast, in RCloud's notebook interface, the extra cells need
to get deleted later, and there seems to be a mental burden that ``bad stuff
is getting saved.''

One thing to try here is making deletion easier, for example with a
shortcut key to remove the last cell, or a way to keep a few cells and cull
the rest. A far more ambitious solution, suggested by one of our
interviewees, is to implement auditing of data analyses as it existed in
S \cite{Becker:1988:Auditing}, so that the code which did not lead to a
result could be discarded automatically.

The interface is also perceived to be slower, both in terms of reaction
time---RCloud necessarily saves a cell to its repository before executing
it---and because it requires mouse interaction.

One thing to try is tagging and filtering the tree on tags, so that users
can mark their own notebooks or those of others to make them easier to
find.

However, this relies on users making a conscious effort, and it may be
better to rely more on passive metrics such as how often a notebook has
been run.

One solution we are mulling over is to allow overwriting the current state
of a notebook with the current state of another fork of the notebook. More
general solutions would be serious engineering endeavors: even GitHub's web
interface does not support merging where there are any conflicts. Actual
sharing of notebooks would probably require moving toward git's notion of
decentralized code repositories, because GitHub gists do not support shared
ownership.

@@

>>>>

@@

{\bf Exploration is hard.}
R is a popular language for exploratory data analysis, and we built RCloud
as a way for data scientists to seamlessly transition from EDA to sharing
and presentation. But the combined notebook and command line interface we designed interferes with the exploration aspect.

In existing R tools, the command prompt is used to try things out. It is very
fast and there is no commitment: nothing gets saved and there is nothing to
later clean up. In contrast, in RCloud's notebook interface, the extra cells need
to get deleted later, and there seems to be a mental burden that \textit{bad stuff
is getting saved}.

One thing to try here is making deletion easier, for example with a
shortcut key to remove the last cell, or a way to keep a few cells and cull
the rest. A far more ambitious solution, suggested by one of our
interviewees, is to implement auditing of data analyses as it existed in S
\cite{Becker:1988:Auditing}, so that a result could be selected and code
which did not lead to it could be discarded automatically.

We got here from an insistence on reproducibility: it is a core principle
of RCloud that no command should be run without being saved, even if it
gets deleted later. This is why we did not implement a simple command line
as many users request. We could imagine making an exception for commands
that have no side-effects, but in the general case this is no easier than
auditing because R is not a purely functional language.

The interface is also perceived to be slower, both for computer and for
human. First, there is a slower reaction to commands being run, because
RCloud necessarily saves a cell to its repository before executing it. It
may be possible to safely write to the repository behind execution of the
same code. Second, and perhaps more detrimental, the interface requires
mouse interaction and switching between the keyboard and mouse, which are
known to be slower. \gordon{we don't need to back that up, right? i hate
  using passive voice.} Shortcut keys can be devised if people are willing
to learn them, but it would be better to reduce the number of such
operations needed.

One thing to try is tagging and filtering the tree on tags, so that users
can mark their own notebooks or those of others to make them easier to
find. However, this relies on users making a conscious effort, and it may be
better to rely more on passive metrics such as how often a notebook has
been run.

One solution we are considering is to allow overwriting the current state
of a notebook with the current state of another fork of the notebook.

More general solutions would be serious engineering endeavors. Even
GitHub's generous web interface does not support merging where there are
conflicts. Actual sharing of notebooks would probably require moving toward
git's notion of decentralized code repositories, because GitHub gists do
not support shared ownership.
